---
title: "Lab 06"
author: "Sylvia Baeyens"
date: "8/1/2021"
output:
  github_document:
  html_document:
    html_preview: false
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo= FALSE, include= FALSE}
#including necessary libraries
library(tidyverse)
library(tidytext)
library(dplyr)
library(ggplot2)
library(data.table)
```

Reading in Data 
```{r}
rawData= "mtsamples.csv"
if(!file.exists(rawData))
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/00_mtsamples/mtsamples.csv", destfile = rawData)

mtsamples= read.csv(rawData) 
mtsamples= as_tibble(mtsamples)
```

## 1. What specialties do we have?

First, we count the number of specialties
```{r}
Nspecialties = mtsamples %>%
  count(medical_specialty)

Nspecialties %>%
  arrange(desc(n)) %>%
  top_n(15) %>%
  knitr::kable()
```

There are `r nrow(Nspecialties)` specialties.

Next, we look at distribution:
```{r}
#using method 2 from class:

ggplot(Nspecialties, aes(x = n, y = fct_reorder(medical_specialty, n))) +
  geom_col()
```

There is no even distribution among the `r nrow(Nspecialties)` specialties. Surgery has the highest count, while hospice-pallative care has the lowest count.

## 2. Tokenizing the words in the transcription column
